DeepSphere: Efficient spherical Convolutional Neural Network with HEALPix sampling for cosmological applications

### 2019-03-04 rebuttal.tex

Indeed. We added a word "typically" to reflect that there may be cases like this.

\begin{mdframed}[style=comment]
For each class, we generate 30 simulations.

- Just mention that there are two classes again, to make it clear that "class" means "cosmology".
\end{mdframed}

% \nati{@Tomek, do you think it is fine that we do not say that class==comoslogy? It think so.}
% \mdeff{I think it's two cosmologies as it's two sets of parameters. Adding again two classes is enough.}

@@

- I think that you should introduce the Gaussian noise paragraph by stating that you perform data augmentation via the addition of randomly drawn Gaussian noise, rather than talking about the random noise first and then explaining that the reason is for data augmentation. ML specialists in particular will be happier with the addition of noise for DA (as will cosmologists I think).
\end{mdframed}

% \nati{I think that the primary goal of adding noise is the make the problem harder, and show robust prediction, not the make the dataset larger... Hence I disagree with the reviewer. What do you think?}
% \TK{This is an interesting observation. The noise is introduced to make the case realistic, as the mass map data in reality is always very noisy. It is convenient that this fact can be used for data augmentation too and make the ML problem easier to solve. As these two facts are connected, we prefer to leave the sentence as is.}

% The primary goal of the noise addition is the make the problem harder and show robust prediction. We use it to augment our dataset as well. Hence we believe that the correct formulation is suitable.

This is an interesting observation. The noise is introduced to make the case realistic, as the mass map data in reality is always very noisy.
It is convenient that this fact can be used for data augmentation too. % and make the ML problem solvable even with a small amount of simulations.
As these two facts are connected, we prefer to leave the sentence as is.

@@

\begin{mdframed}[style=comment]
Probabilities are computed by the SM layer.

- As explained above, these are not probabilities, they are class predictions from the network.
\end{mdframed}

% \nati{Should we simply say "normalized network score"? I do not know...}
% \mdeff{SM normalizes the predictions}

We updated the sentence.

@@

- What is "it" here? DeepSphere or graph networks or something else?
\end{mdframed}

"it" refer to "graphs" in general. We have updated the manuscript.

\begin{mdframed}[style=comment]
We believe that it might be the case for a different construction of the graph (fully connected) and that a proof could be built on top of the work of [33]. That is, however, out of the scope of this contribution.

- What makes you believe this? If you don't want to mention it in this paper, please don't speculate a "belief". I might just be reading the sentence incorrectly, but it makes me a little uncomfortable when there is no evidence backing that up. If you have some evidence then it's worth presenting it.
\end{mdframed}

As already mentioned in this rebuttal, our belief comes from the fact that we are working on this.
Nevertheless, we agree that a scientific article should not contain speculations.
We removed that speculation.

@@

- There is a superfluous word (dealing/working). I think this would read better as "Nevertheless, when working with only a part of the sphere, the border slightly affects the convolution."
\end{mdframed}

We corrected the wording.

\begin{mdframed}[style=comment]
These border effects can, however, be mitigated by padding with zeros a small area around the part of interest.

- Are they really mitigated? Surely there are still effects but you are happy with the level?
\end{mdframed}

With the zero padding, the border effects become similar to the ones in traditional CNNs.
This information has been added to the manuscript.

### 2019-03-04 scnn.tex
comment 2.14 (details) part 4

As discussed in \secref{architecture}, the choice of a network architecture depends on the data and task at hand.
For our problem, where we assume that each pixel carries the same quantity of information about the cosmological model, a rotation invariant FCN architecture should be best.
The selected network is defined as
\begin{equation*}
	FCN = \underbrace{SM \circ AV \circ GC_2}_{\text{classifier}} \circ \underbrace{L_5 \circ L_4 \circ L_3 \circ L_2 \circ L_1}_{\text{feature extractor}},
\end{equation*}
\begin{center}
with \hspace{0.5cm} \begin{tabular}{lll}
   $L_1$ &  $=$ &$ P_4 \circ \sigma \circ BN \circ GC_{16}$, \\
   $L_2$ &  $=$ &$ P_4 \circ \sigma \circ BN \circ GC_{32}$, \\
   $L_3,L_4,L_5$ & $=$ &$ P_4 \circ \sigma \circ BN \circ GC_{64}$, \\
\end{tabular}
\end{center}
where $GC_F$ indicates a graph convolutional layer with $F$ output feature maps, $P_4$ is a pooling layer that divides the number of pixels by 4, $\sigma$ is a ReLU, and the Chebyshev polynomials in \eqnref{graph_convolution_cheby} are of degree $K=5$.
The layers $L_1$ to $L_5$ build the statistical evidence to classify each pixel of a down-sampled sphere. The last $GC$ layer acts as a linear classifier and outputs two (non-normalized) predictions. The $AV$ layer averages these predictions. The normalized network scores are computed by the $SM$ layer.

@@

While testing the FCN and CNN variants, we made the following empirical observations.
First, training was more stable (in the sense that the loss decreased more smoothly) when using Chebyshev polynomials, as in \eqnref{graph_convolution_cheby}, rather than monomials, as in \eqnref{graph_convolution_monomial}.
Nevertheless, we could not observe a significant difference in accuracy after convergence.
Second, using $\ell_2$ regularization does not help either with performance or training of the models, mostly because the models are not over-fitting.
Third, we recommend initializing the Chebyshev coefficients with a centered Gaussian distribution with standard deviation $\sqrt{2/(F_{in} \times (K + 0.5))}$, where $K$ is the degree of the Chebyshev polynomials and $F_{in}$ the number of input channels. This standard deviation has the property of keeping the energy of the signal more or less constant across layers\footnote{We derived this rule from the Chebyshev polynomial equations and some empirical experiments. \nati{Following the principle presented in \cite{glorot2010understanding}, we searched for the initialisation scheme keeping the expected amplitude of the signal constant through the different layers of the neural network.}}.
% Nati to Michael: The problem with the previous claim is that it depends on the non-linarity. Furthermore, does it make sense to make such a claim if we use batch norm?
Finally, we observe that scaling the Laplacian's eigenvalues between $[-a, a]$, where $0 < a \leq 1$, significantly helps in stabilizing the optimization. We use $a = 0.75$ in our experiments.

@@

While these indications show that the constructed graph Fourier basis approximates well the spherical harmonics, one should not forget that the small irregularities in the sampling (non-constant number of neighboring pixels and varying distance between pixels, see \figref{healpix_graph_4}) have an important effect on the graph Fourier modes.
First, we believe that they are responsible for energy leaking across frequency bands in \figref{subspace_harmonics_eigenvectors}.
Second, as the resolution is increased with $N_{side} \rightarrow \infty$ and $N_{pix} \rightarrow \infty$, we are still unsure if the eigenvectors would converge towards the spherical harmonics.
% We believe that it might be the case for a different construction of the graph (fully connected) and that a proof could be built on top of the work of~\cite{belkin2007convergence}. That is, however, out of the scope of this contribution.
The theoretical study of the convergence is, however, out of the scope of this contribution.
Third, counter-intuitively, some eigenvectors will be localized~\citep{perraudin2018global}, i.e., they will span a small part of the sphere.
Those discrepancies result in a convolution operation that is not exactly equivariant to rotation.
Nevertheless, our experiments suggest that these downsides do not have an important effect on the convolution nor hinder the performance of the neural network.

@@

\begin{figure}[ht!]
  \centering
  \includegraphics[width=\linewidth]{gaussian_filters_sphere}
  \caption{Comparison of convolution with the graph and the spherical harmonics ($N_{side} = 16$).
  Top: Diffusion of a unit of heat for different times $t$ using the graph.
  Bottom: spherical symmetric Gaussian smoothing for different $\sigma$ (arcmin).
  Relative difference between graph convolution and spherical smoothing: $10.4$\%, $4.8$\%, $3.8$\%.
}
  \label{fig:gaussian_filters_comparizon}
\end{figure}
Furthermore, the harmonic resemblance is another sign that the constructed graph is able to capture the spherical
structure of the HEALPix sampling. In some applications, where the
exactitude of the convolution is not a requirement, such as de-noising, graph convolution could be used instead of using the spherical harmonics.
In a neural network setting, the fact that the graph convolution is not exactly equivariant to rotation is probably compensated by the fact that convolution kernels are learned.

@@

The graph setting used throughout this contribution corresponds to assuming reflective border conditions.
While that is irrelevant when working on the complete sphere (as it has no border), it slightly affects the convolution operation when only a part of the sphere is considered.
As depicted in \figref{border_effects}, a filter localized near a border (via $h(\L) \b \delta_i$) is no longer isotropic.
These border effects can, however, be mitigated by padding with zeros a small area around the part of interest.
In this case, they are similar to the border effects in traditional CNNs.
We however do not expect these effects to cause any problem as long as the data samples cover the same area.

>>>>

As discussed in \secref{architecture}, the choice of a network architecture depends on the data and task at hand.
For our problem, where we assume that each pixel carries the same quantity of information about the cosmological model, a rotation invariant FCN architecture should be best.
The selected network is defined as
\begin{equation*}
	FCN = \underbrace{SM \circ AV \circ GC_2}_{\text{classifier}} \circ \underbrace{L_5 \circ L_4 \circ L_3 \circ L_2 \circ L_1}_{\text{feature extractor}},
\end{equation*}
\begin{center}
with \hspace{0.5cm} \begin{tabular}{lll}
   $L_1$ &  $=$ &$ P_4 \circ \sigma \circ BN \circ GC_{16}$, \\
   $L_2$ &  $=$ &$ P_4 \circ \sigma \circ BN \circ GC_{32}$, \\
   $L_3,L_4,L_5$ & $=$ &$ P_4 \circ \sigma \circ BN \circ GC_{64}$, \\
\end{tabular}
\end{center}
where $GC_F$ indicates a graph convolutional layer with $F$ output feature maps, $P_4$ is a pooling layer that divides the number of pixels by 4, $\sigma$ is a ReLU, and the Chebyshev polynomials in \eqnref{graph_convolution_cheby} are of degree $K=5$.
The layers $L_1$ to $L_5$ build the statistical evidence to classify each pixel of a down-sampled sphere.
The last $GC$ layer acts as a linear classifier and outputs two predictions.
The $AV$ layer averages these predictions and the $SM$ layer normalizes them.

@@

While testing the FCN and CNN variants, we made the following empirical observations.
First, training was more stable (in the sense that the loss decreased more smoothly) when using Chebyshev polynomials, as in \eqnref{graph_convolution_cheby}, rather than monomials, as in \eqnref{graph_convolution_monomial}.
Nevertheless, we could not observe a significant difference in accuracy after convergence.
Second, using $\ell_2$ regularization does not help either with performance or training of the models, mostly because the models are not over-fitting.
Third, we recommend initializing the Chebyshev coefficients with a centered Gaussian distribution with standard deviation $\sqrt{2/(F_{in} \times (K + 0.5))}$, where $K$ is the degree of the Chebyshev polynomials and $F_{in}$ the number of input channels.
This standard deviation has the property of keeping the energy of the signal more or less constant across layers.\footnote{We derived this rule from the Chebyshev polynomial equations, the principles presented in \cite{glorot2010understanding}, and some empirical experiments.}
% Nati to Michael: The problem with the previous claim is that it depends on the non-linarity. Furthermore, does it make sense to make such a claim if we use batch norm?
% MD: I don't know... But that's not too important.
Finally, we observe that scaling the Laplacian's eigenvalues between $[-a, a]$, where $0 < a \leq 1$, significantly helps in stabilizing the optimization. We use $a = 0.75$ in our experiments.

@@

While these indications show that the constructed graph Fourier basis approximates well the spherical harmonics, one should not forget that the small irregularities in the sampling (non-constant number of neighboring pixels and varying distance between pixels, see \figref{healpix_graph_4}) have an important effect on the graph Fourier modes.
First, we believe that they are responsible for energy leaking across frequency bands in \figref{subspace_harmonics_eigenvectors}.
Second, as the resolution is increased with $N_{side} \rightarrow \infty$ and $N_{pix} \rightarrow \infty$, we are still unsure if the eigenvectors would converge towards the spherical harmonics.
% We believe that it might be the case for a different construction of the graph (fully connected) and that a proof could be built on top of the work of~\cite{belkin2007convergence}. That is, however, out of the scope of this contribution.
The theoretical study of those phenomenons is left as future work.
Third, counter-intuitively, some eigenvectors will be localized~\citep{perraudin2018global}, i.e., they will span a small part of the sphere.
Those discrepancies result in a convolution operation that is not exactly equivariant to rotation.
Nevertheless, our experiments suggest that these downsides do not have an important effect on the convolution nor hinder the performance of the neural network.

@@

\begin{figure}[ht!]
  \centering
  \includegraphics[width=\linewidth]{gaussian_filters_sphere}
  \caption{Comparison of convolution with the graph and the spherical harmonics ($N_{side} = 16$).
  Top: Diffusion of a unit of heat for different times $t$ using the graph.
  Bottom: spherical symmetric Gaussian smoothing for different $\sigma$ (arcmin).
  Relative difference between graph convolution and spherical smoothing: $10.4$\%, $4.8$\%, $3.8$\%.
}
  \label{fig:gaussian_filters_comparizon}
\end{figure}

Furthermore, the harmonic resemblance is another sign that the constructed graph is able to capture the spherical structure of the HEALPix sampling.
In some applications, where the exactitude of the convolution is not a requirement, such as de-noising, graph convolution could be used instead of using the spherical harmonics.
In a neural network setting, the fact that the graph convolution is not exactly equivariant to rotation is probably compensated by the fact that convolution kernels are learned.

@@

The graph setting used throughout this contribution corresponds to assumed reflective border conditions.
While that is irrelevant when working on the complete sphere (as it has no border), it slightly affects the convolution operation when only a part of the sphere is considered.
As depicted in \figref{border_effects}, a filter localized near a border (via $h(\L) \b \delta_i$) is no longer isotropic.
These border effects can, however, be mitigated by padding with zeros a small area around the part of interest (in which case they become similar to border effects in classical CNNs).
We however do not expect these effects to cause any problem as long as the data samples cover the same area.
